---
title: "Assignment 1"
author: "Alessandro Kerr"
date: " Sept 26, 2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F, message = F)

library(tidyverse)
library(plotly)
library(sf)
library(tigris)
library(leaflet)
library(censusapi)

pge_elec_all <- NULL

```

```{r}
year <- 2017:2020
quarters <- 1:4
type <- "Electric"

pge_elec <- NULL

for (quarter in quarters) {
  filename <- 
    paste0("PGE_",
         year,
         "_Q", 
         quarter,
         "_",
         type,
         "UsageByZip.csv")
  
  temp <- read_csv(filename)
  pge_elec <- rbind(pge_elec, temp)
  saveRDS(pge_elec, "pge_elec")
  
  
}
```


```{r}
year <- 2021
quarters <- 1:2
type <- "Electric"

pge_elec2 <- NULL

for (quarter in quarters) {
  filename <- 
    paste0("PGE_",
         year,
         "_Q", 
         quarter,
         "_",
         type,
         "UsageByZip.csv")
  
  temp <- read_csv(filename)
  pge_elec2 <- rbind(pge_elec2, temp)
  saveRDS(pge_elec2, "pge_elec2")
  
  
}

pge_elec_all <- rbind(pge_elec, pge_elec2)

```

ˆ
```{r}
year <- 2017:2020
quarters <- 1:4
type <- "Gas"

pge_gas <- NULL

for (quarter in quarters) {
  filename <- 
    paste0("PGE_",
         year,
         "_Q", 
         quarter,
         "_",
         type,
         "UsageByZip.csv")
  
  temp <- read_csv(filename)
  pge_gas <- rbind(pge_gas, temp)
  saveRDS(pge_gas, "pge_gas")
}
```


```{r}
year <- 2021
quarters <- 1:2
type <- "Gas"

pge_gas2 <- NULL

for (quarter in quarters) {
  filename <- 
    paste0("PGE_",
         year,
         "_Q", 
         quarter,
         "_",
         type,
         "UsageByZip.csv")
  
  temp <- read_csv(filename)
  pge_gas2 <- rbind(pge_gas2, temp)
  saveRDS(pge_gas2, "pge_gas")
}

pge_gas_all <- rbind(pge_gas, pge_gas2)

```

```{r}

pge_filter <- 
  filter(
  pge_elec_all,
  CUSTOMERCLASS ==
    "Elec- Commercial")

pge_select <-
  select(
    pge_filter,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TOTALKWH
  )

pge_mutate<-
  mutate(
    pge_select,
    TOTALKBTU = 
      TOTALKWH*3.412
      )

pge_select <- 
  select(
    pge_mutate,
    -TOTALKWH
  )

pge_filter2 <- 
  filter(
  pge_elec_all,
  CUSTOMERCLASS ==
    "Elec- Residential")

pge_select2 <-
  select(
    pge_filter2,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TOTALKWH
  )

pge_mutate2<-
  mutate(
    pge_select2,
    TOTALKBTU = 
      TOTALKWH*3.412
      )

pge_select2 <- 
  select(
    pge_mutate2,
    -TOTALKWH
  )

# Now for Gas

pge_filter3 <- 
  filter(
  pge_gas_all,
  CUSTOMERCLASS ==
    "Gas- Commercial")

pge_select3 <-
  select(
    pge_filter3,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TOTALTHM
  )

pge_mutate3<-
  mutate(
    pge_select3,
    TOTALKBTU = 
      TOTALTHM*99.976
      )

pge_select3 <- 
  select(
    pge_mutate3,
    -TOTALTHM
  )

pge_filter4 <- 
  filter(
  pge_gas_all,
  CUSTOMERCLASS ==
    "Gas- Residential")

pge_select4 <-
  select(
    pge_filter4,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TOTALTHM
  )

pge_mutate4<-
  mutate(
    pge_select4,
    TOTALKBTU = 
      TOTALTHM*99.976
      )

pge_select4 <- 
  select(
    pge_mutate4,
    -TOTALTHM
  )


```

```{r}
# numbers <- 1:4
# for (number in numbers) {
#   file <- paste0(
#     "pge_select",
#     number)
# 

DATE = NULL

  pge_mutate <- 
    mutate(
      pge_select,
      DATE =
        paste(
          pge_select$YEAR,
          pge_select$MONTH,
          "01",
          sep="-"
        ) %>% as.Date()
    )
  
  
  pge_mutate2 <- 
    mutate(
      pge_select2,
      DATE =
        paste(
          pge_select2$YEAR,
          pge_select2$MONTH,
          "01",
          sep="-"
        ) %>% as.Date()
    )
  
  pge_mutate3 <- 
    mutate(
      pge_select3,
      DATE =
        paste(
          pge_select3$YEAR,
          pge_select3$MONTH,
          "01",
          sep="-"
        ) %>% as.Date()
    )
  
  
  pge_mutate4 <- 
    mutate(
      pge_select4,
      DATE =
        paste(
          pge_select4$YEAR,
          pge_select4$MONTH,
          "01",
          sep="-"
        ) %>% as.Date()
    )
  
  pge_final_com = NULL
  pge_final_res = NULL
   
  
  pge_final_com <- rbind(pge_mutate, pge_mutate3)
  pge_final_res <- rbind(pge_mutate2, pge_mutate4)
  
  
  
# }
 

```

```{r}
pge_chart_com <- 
  pge_final_com %>% 
  ggplot() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y") +
  geom_bar( 
    aes(
      x = DATE ,
      y = TOTALKBTU,
      fill = CUSTOMERCLASS,
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Month",
    y = "kBTU's",
    title = "PG&E Territory Monthly Commercial Electricity and Gas Usage, 2017 to Present",
    fill = "Utility Type"
  )

  
pge_chart_com

```

There are noticable differences between the Commerical utility use, notably electricity, during the COVID pandemic. In April 2020, a sharp decrease in commerical electricity use can be seen on the above chart. 

```{r}
pge_chart_res <- 
  pge_final_res %>% 
  ggplot() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y") +
  geom_bar( 
    aes(
      x = DATE,
      y = TOTALKBTU,
      fill = CUSTOMERCLASS,
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Month",
    y = "kBTU's",
    title = "PG&E Territory Monthly Residential Electricity and Gas Usage, 2017 to Present",
    fill = "Utility Type"
  )

  
pge_chart_res
```

In terms of residential electricity and gas use across the four analysis years, there was a slight increase as a result of the pandemic. This is reasonable as there were more people working and spending extra time at home. 

Key assumptions: It is assumed that the given CSV files obtained from the Pacific Gas and Electricity Company are accurate and representative of the actual trends.   
