---
title: "Week8Chapter8Class2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(censusapi)
library(StatMatch)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  )

bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

bay_tracts <-
  tracts("CA", bay_county_names, cb = T, progress_bar = F)

bay_multiple_tract <- 
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "tract:*",
    regionin = "state:06+county:001,013,041,055,075,081,085,095,097",
    vars = c(
      "B06009_001E",
      "B06009_002E",
      "B06009_003E",
      "B19001_001E",
      "B19001_014E",
      "B19001_015E",
      "B19001_016E",
      "B19001_017E",
      "B19001A_001E"
    )
  ) %>% 
  transmute(
    tract = paste0(state, county, tract),
    perc_college = 1 - (B06009_002E + B06009_003E) / B06009_001E,
    perc_over100k = (B19001_014E + B19001_015E + B19001_016E + B19001_017E) / B19001_001E,
    perc_white = B19001A_001E / B19001_001E
  ) %>% 
  filter(
    !is.na(perc_college), 
    !is.na(perc_over100k),
    !is.na(perc_white)
  )
```

```{r}
obs_matrix <-
  bay_multiple_tract %>% 
  select(
    perc_white, 
    perc_over100k,
    perc_college
  ) %>% 
  as.matrix()
```

```{r}
dist_matrix <- mahalanobis.dist(obs_matrix)

rownames(dist_matrix) <- bay_multiple_tract$tract
colnames(dist_matrix) <- bay_multiple_tract$tract
```

```{r}
dist_matrix_pairmatch <- dist_matrix
diag(dist_matrix_pairmatch) <- NA

matched_pair_tract <-
  1:nrow(dist_matrix_pairmatch) %>% 
  map_dfr(function(x){
    
    min_index <- which(dist_matrix_pairmatch[x, ] == min(dist_matrix_pairmatch[x, ], na.rm = T))
   
    data.frame(
      tract = bay_multiple_tract$tract[x],
      matched_tract = bay_multiple_tract$tract[min_index]
    )
    
  })
```

```{r}
leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = bay_tracts %>% 
      filter(GEOID == matched_pair_tract[2,1])
  ) %>% 
  addPolygons(
    data = bay_tracts %>% 
      filter(GEOID == matched_pair_tract[2,2])
  ) %>% 
  addPolygons(
    data = bay_tracts %>% 
      filter(GEOID == matched_pair_tract[3,1]),
    color = "green"
  ) %>% 
  addPolygons(
    data = bay_tracts %>% 
      filter(GEOID == matched_pair_tract[3,2]),
    color = "green"
  ) %>% 
  addPolygons(
    data = bay_tracts %>% 
      filter(GEOID == matched_pair_tract[4,1]),
    color = "red"
  ) %>% 
  addPolygons(
    data = bay_tracts %>% 
      filter(GEOID == matched_pair_tract[4,2]),
    color = "red"
  )
```

```{r}
match_set_tract <- dist_matrix["06081611900", ] %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(
    tract = rowname,
    match = "."
  ) %>% 
  right_join(
    bay_multiple_tract
  ) %>% 
  arrange(match) %>% 
  .[1:21, ] %>% 
  left_join(bay_tracts %>% select(tract = GEOID)) %>% 
  st_as_sf()

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = match_set_tract[1, ],
    color = "red"
  ) %>% 
  addPolygons(
    data = match_set_tract[-1, ]
  )
```

#6.2

```{r}
dataframe %>% st_as_sf(coords = c("lng_name","lat_name"), crs = 4326)
#coords to sf object

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = bay_pumas,
    weight = 1,
    color = "gray",
    label = ~PUMACE10
  ) %>% 
  addMarkers(
    lng = -121.9415017,
    lat = 37.502171
  ) %>% 
  addPolygons(
    data = bay_pumas %>% 
      filter(PUMACE10 == "08504")
  )
```

```{r}
pums_2014_2019 <- readRDS("pums_2014_2019_wts.rds")
  
pums_bart <- pums_2014_2019 %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    bart = ifelse(
      JWTR %in% c("4"),
      PWGTP,
      0
    )
  ) %>% 
  group_by(PUMA, year) %>% 
  summarize(
    pop = sum(PWGTP),
    bart = sum(bart)
  )
```

```{r}
pums_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = pums_bart %>% 
    filter(year == 2017) %>% 
    pull(pop)
)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = pums_bart %>% 
      filter(year == 2017) %>% 
      right_join(bay_pumas %>% select(PUMA = PUMACE10)) %>% 
      st_as_sf(),
    fillColor = ~pums_pal(pop),
    color = "white",
    weight = 1,
    fillOpacity = 0.5,
    label = ~paste0(PUMA,": Population ", pop)
  )
```

```{r}
pums_pal <- colorNumeric(
  palette = "GnBu",
  domain = pums_bart %>% 
    filter(year == 2017) %>% 
    pull(bart)
)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = pums_bart %>% 
      filter(year == 2017) %>% 
      right_join(bay_pumas %>% select(PUMA = PUMACE10)) %>% 
      st_as_sf(),
    fillColor = ~pums_pal(bart),
    color = "white",
    weight = 1,
    fillOpacity = 0.5,
    label = ~paste0(PUMA,": ", bart, " BART commute riders")
  )
```

```{r}
pums_bart_clean <-
  pums_bart %>% 
  select(-pop) %>% 
  pivot_wider(
    names_from = year,
    values_from = bart
  ) 

#pivot wider if need to do math across several years
```

```{r}
obs_matrix <-
  pums_bart_clean %>% 
  ungroup() %>% 
  select(`2014`,`2015`,`2016`) %>% 
  as.matrix()

dist_matrix <- mahalanobis.dist(obs_matrix)

rownames(dist_matrix) <- pums_bart_clean$PUMA
colnames(dist_matrix) <- pums_bart_clean$PUMA

match <- dist_matrix["08504",] %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(
    PUMA = rowname,
    match = "."
  ) %>% 
  right_join(
    pums_bart_clean
  ) %>% 
  arrange(match) %>% 
  .[1:11, ] %>% 
  left_join(bay_pumas %>% select(PUMA = PUMACE10)) %>% 
  st_as_sf()
```

```{r}
leaflet() %>% 
  addTiles() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = match[1, ],
    color = "red",
    label = ~PUMA
  ) %>% 
  addPolygons(
    data = match[-1, ],
    label = ~PUMA
  )
```

```{r}
match_pumas <-
  match %>% 
  filter(!PUMA %in% c("08504")) %>% 
  st_drop_geometry() %>% 
  select(-match) %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  ) %>%
  group_by(
    year
  ) %>% 
  summarize(
    bart = mean(bart),
    PUMA = "Similar PUMAs"
  )

treatment_pumas <-
  match %>% 
  filter(PUMA %in% c("08504")) %>% 
  select(-match) %>% 
  st_drop_geometry() %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  )

rbind(
  treatment_pumas,
  match_pumas
) %>% 
  ggplot(
    aes(
      x = as.numeric(year),
      y = bart,
      color = PUMA
    )
  ) +
  geom_line() +
  geom_vline(xintercept = 2017, linetype = "dashed") +
  labs(
    title = "Milpitas vs. control neighborhoods, BART ridership",
    x = "Year",
    y = "BART commute riders"
  )
```


#treated: minor difference between treated and control group before bart station
#time: increases by 200 bart riders per year (not effect of bart station - happened to control group)
# treated:time is the different between red and green before and after 2017 (due to bart station at a parrticular point in time affecting particular puma). (in if statement - only returns results that were in that puma and beyond 2017 - so isolates effect of bart station and time)



#brute force method? 8 simulations? for 7 point