---
title: "AlexanderNg_Assignment4"
author: "Alexander Ng"
date: "3/1/2022"
output: html_document
---

```{r}
library(raster)
library(stringr)
library(tigris)
library(tidyverse)
library(readxl)
library(censusapi)
library(sf)
library(leaflet)
library(htmltools)
library(tidycensus)
library(mapview)
library(dplyr)
library(ggrepel)
library(ggplot2)
library(jsonlite)
library(stars)
library(mapboxapi)
library(tiff)

knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F, message = F)

```

## PART 1 - Flood Maps for Foster City
```{r}
#San Mateo Zips & Blocks

# zctas <- zctas()
# 
# zip <- zctas %>% 
#   filter(GEOID10 == "94404")
# 
# saveRDS(zip, "zip.rds")

# zip <- readRDS("zip.rds")
# 
# zip_blocks <-
#   blocks("CA","San Mateo", progress_bar = F, year = 2020) 
# zip_blocks <- zip_blocks %>%
#   .[zip,]
# 
# saveRDS(zip_blocks, "zip_blocks.rds")
  
```

Foster City Boundary
```{r}
mapview(zip)
```
#use OCOS 50cm slp 100 year rp to determine flood-affected region

```{r}

# for(slr in c("000","025","050")){
#   
#   for(rp in c("001","020","100")){
#     
#     print(paste0("SLR",slr,"_RP",rp))
#     
#     path <- paste0("/Volumes/GoogleDrive-112836525701087151759/Shared drives/SFBI/Data Library/OCOF/san_mateo_flooding_slr",slr,"/flooding/v2.1/county_san_mateo_flddepth_slr",slr,"_w",rp,".tif")
#     
#     flood <- raster(path) %>% 
#       crop(
#         zip %>% 
#           st_transform(26910) %>% 
#           st_bbox()
#       )
#   
#     writeRaster(flood,paste0("/Users/xander/Documents/GitHub/alexngld.github.io/flood/SLR",slr,"_RP",rp,"_fc_flood.tif"), overwrite = T)
#     
#   }
# }


flood_SLR000_RP001<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR000_RP001_fc_flood.tif")
flood_SLR000_RP020<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR000_RP020_fc_flood.tif")
flood_SLR000_RP100<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR000_RP100_fc_flood.tif")
flood_SLR025_RP001<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR025_RP001_fc_flood.tif")
flood_SLR025_RP020<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR025_RP020_fc_flood.tif")
flood_SLR025_RP100<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR025_RP100_fc_flood.tif")
flood_SLR050_RP001<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR050_RP001_fc_flood.tif")
flood_SLR050_RP020<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR050_RP020_fc_flood.tif")
flood_SLR050_RP100<-raster("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR050_RP100_fc_flood.tif")

```

```{r}

flood_pal3 <- colorNumeric(
  palette = "Blues",
  domain = values(flood_SLR000_RP100),
  na.color = "transparent"
)

flood_pal6 <- colorNumeric(
  palette = "Blues",
  domain = values(flood_SLR025_RP100),
  na.color = "transparent"
)

flood_pal9 <- colorNumeric(
  palette = "Blues",
  domain = values(flood_SLR050_RP100),
  na.color = "transparent"
)

```

SLR000
```{r}
leaflet() %>% 
  addMapboxTiles(
    style_id = "satellite-streets-v11",
    username = "mapbox",
    options = tileOptions(opacity = 0.5)
  ) %>% 
  addRasterImage(
    flood_SLR000_RP001,
    colors = flood_pal3,
    group = "flood_SLR000_RP001"
  ) %>% 
  addRasterImage(
    flood_SLR000_RP020,
    colors = flood_pal3,
    group = "flood_SLR000_RP020"
  ) %>%
    addRasterImage(
    flood_SLR000_RP100,
    colors = flood_pal3,
    group = "flood_SLR000_RP100"
  ) %>% 
    addLegend(
    pal = flood_pal3,
    values = values(flood_SLR000_RP100),
    title = "Flood depth, cm"
  ) %>%
  addLayersControl(
    baseGroups = c("flood_SLR000_RP001","flood_SLR000_RP020","flood_SLR000_RP100"),
    options = layersControlOptions(collapsed = FALSE)
    )
```

SLR025
```{r}
leaflet() %>% 
  addMapboxTiles(
    style_id = "satellite-streets-v11",
    username = "mapbox",
    options = tileOptions(opacity = 0.5)
  ) %>% 
  addRasterImage(
    flood_SLR025_RP001,
    colors = flood_pal6,
    group = "flood_SLR025_RP001"
  ) %>% 
  addRasterImage(
    flood_SLR025_RP020,
    colors = flood_pal6,
    group = "flood_SLR025_RP020"
  ) %>%
    addRasterImage(
    flood_SLR025_RP100,
    colors = flood_pal6,
    group = "flood_SLR025_RP100",
    maxBytes = 5*1024*1024
  ) %>%
    addLegend(
    pal = flood_pal6,
    values = values(flood_SLR025_RP100),
    title = "Flood depth, cm"
  ) %>%
  addLayersControl(
     baseGroups = c("flood_SLR025_RP001","flood_SLR025_RP020","flood_SLR025_RP100"),
    options = layersControlOptions(collapsed = FALSE)
    )
  
```

SLR050
```{r}
leaflet() %>% 
  addMapboxTiles(
    style_id = "satellite-streets-v11",
    username = "mapbox",
    options = tileOptions(opacity = 0.5)
  ) %>% 
  addRasterImage(
    flood_SLR050_RP001,
    colors = flood_pal9,
    group = "flood_SLR050_RP001"
  ) %>% 
  addRasterImage(
    flood_SLR050_RP020,
    colors = flood_pal9,
    group = "flood_SLR050_RP020"
  ) %>%
    addRasterImage(
    flood_SLR050_RP100,
    colors = flood_pal9,
    group = "flood_SLR050_RP100",
    maxBytes = 5*1024*1024
  ) %>%
    addLegend(
    pal = flood_pal9,
    values = values(flood_SLR050_RP100),
    title = "Flood depth, cm"
  ) %>%
  addLayersControl(
    baseGroups = c("flood_SLR050_RP001","flood_SLR050_RP020","flood_SLR050_RP100"),
    options = layersControlOptions(collapsed = FALSE)
    )
```

## PART 2 - EMFAC Vehicle Counts in Foster City
```{r}
emfac <- 
  read_csv("EMFAC2021-EI-202xClass-SanMateo(SF)2020-2030-2040-2050-Annual-20220218111115.csv", skip = 8) %>%
  group_by(`Calendar Year`)%>%
  summarize(`Vehicle Count` = sum(Population)) %>%
  rbind(NA) %>%
  mutate(Calc=lag(`Vehicle Count`))%>%
  mutate(Increase=((`Vehicle Count`/Calc)-1)*100) %>%
  dplyr::select(`Calendar Year`, Increase) %>%
  filter(!is.na(Increase))

```

```{r}
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables",
  )

# zip_blocks <- blocks %>%
#   st_centroid() %>%
#   .[zip,] %>%
#   st_drop_geometry() %>%
#   left_join(blocks %>% select(GEOID10)) %>%
#   st_as_sf() # blocks within zip code
# 
# saveRDS(zip_blocks, "zip_blocks.rds")
zip_blocks<-readRDS("zip_blocks.rds") %>%
  mutate(cbg = as.character(substr(GEOID20, 1, 12)))


# fc_bg_vehicle <-
#   getCensus(
#     name = "acs/acs5",
#     vintage = 2019,
#     region = "block group:*",
#     regionin = "state:06+county:081",
#     vars = "group(B25044)"
#   ) %>% 
#   mutate(
#     cbg =
#       paste0(state,county,tract,block_group)
#   ) %>% 
#   dplyr::select(!c(GEO_ID,state,county,tract,block_group,NAME) & !ends_with(c("EA","MA","M"))) %>%
#   pivot_longer(
#     ends_with("E"),
#     names_to = "variable",
#     values_to = "estimate"
#   ) %>%
#   left_join(
#     acs_vars_2019_5yr %>% 
#       dplyr::select(name, label), 
#     by = c("variable" = "name")
#   ) %>% 
#   dplyr::select(-variable) %>% 
#   separate(
#     label,
#     into = c(NA,NA,"tenure","vehicle"),
#     sep = "!!"
#   ) %>% 
#   filter(!is.na(vehicle)) %>% 
#   filter(cbg %in% fc_cbg$GEOID)
# 
# fc_no_one_vehicle <- fc_bg_vehicle %>%
#   group_by(cbg, vehicle) %>%
#   summarize(estimate = sum(estimate)) %>%
#   filter(vehicle == "No vehicle available" | vehicle == "1 vehicle available")
# 
# fc_vehicles <- fc_bg_vehicle %>% 
#   filter(vehicle != "No vehicle available") %>% 
#   mutate(
#     vehicles = substr(vehicle, 1,1) %>% as.numeric(),
#     vehicle_count = vehicles * estimate
#   ) %>% 
#   group_by(cbg) %>% 
#   summarize(estimate = sum(estimate))
#saveRDS(fc_vehicles, "fc_vehicles.rds")


fc_vehicles <- readRDS("fc_vehicles.rds")

```

## PART 3 - Foster City Flood Exposure by Census Block
```{r}
leaflet() %>% 
  addMapboxTiles(
    style_id = "streets-v11",
    username = "mapbox"
  ) %>% 
  addPolygons(
    data = zip_blocks,
    fill = F, 
    weight = 1
  ) %>%
  addPolygons(
    data= zip,
    fill = F,
    weight = 2,
    color = "red"
  )

```


```{r}
# dec_vars_2020 <-
#   listCensusMetadata(
#     name = "2020/dec/pl",
#     type = "variables"
#   )

# fc_pop_2020 <-
#   getCensus(
#     name = "dec/pl",
#     vintage = 2020,
#     region = "block:*",
#     regionin = "state:06+county:081",
#     vars = "group(P1)"
#   ) %>%
#   mutate(
#     block =
#       paste0(state,county,tract,block)
#   ) %>%
#   filter(block %in% zip_blocks$GEOID20) %>%
#   select(!c(GEO_ID,state,county,tract,NAME) & !ends_with(c("NA"))) %>%
#   pivot_longer(
#     ends_with("N"),
#     names_to = "name",
#     values_to = "estimate"
#   ) %>%
#   filter(name=="P1_001N") %>%
#   left_join(
#     zip_blocks %>%
#       select("GEOID20"),
#     by = c("block"="GEOID20")
#   ) %>%
#   st_as_sf()

# saveRDS(fc_pop_2020, "fc_pop_2020.rds")

fc_pop_2020<-readRDS("fc_pop_2020.rds")

```

```{r}
# path <- "/Volumes/GoogleDrive-112836525701087151759/Shared drives/SFBI/Data Library/OSM/gis_osm_buildings_a_free_1.shp"
# 
# osm_bldg <- st_read("/Volumes/GoogleDrive-112836525701087151759/Shared drives/SFBI/Data Library/OSM/gis_osm_buildings_a_free_1.shp")
# 
# fc_cbg <- zip_blocks %>% 
#   st_transform(st_crs(osm_bldg))
# 
# bldg_osm <-
#   osm_bldg[fc_cbg, ]
# 
# saveRDS(bldg_osm, "bldg_osm.rds")

```

```{r}
bldg_osm<-readRDS("bldg_osm.rds") %>%
  st_transform(st_crs(4269))

fc_bldg <- bldg_osm[zip, ] 

mapview(fc_bldg)

```

```{r}
# mapview(fc_pop_2020)

mapview(zip_blocks)
```

```{r}
flood_max <- 
  raster("flood/SLR050_RP100_fc_flood.tif")

flood_max_extent <- 
  flood_max %>% 
  st_as_stars() %>% 
  mutate(SLR050_RP100_fc_flood = ifelse(
    !is.na(SLR050_RP100_fc_flood),
    1,
    NA
  )) %>% 
  st_as_sf(merge = T) %>% 
  st_set_crs(26910) %>% 
  st_make_valid() %>% 
  st_transform(4326)

fc_bldg_flooded_max <-
  fc_bldg %>% 
  st_transform(4326) %>% 
  .[flood_max_extent,]

```

```{r}
flood_pal <- colorNumeric(
  palette = "Blues",
  domain = values(flood_max),
  na.color = "transparent"
)

leaflet() %>% 
  addMapboxTiles(
    style_id = "satellite-streets-v11",
    username = "mapbox",
    options = tileOptions(opacity = 0.5)
  ) %>% 
  addRasterImage(
    flood_max,
    colors = flood_pal,
    opacity = 0.75,
    maxBytes = 5*1024*1024
  ) %>% 
  addPolygons(
    data = fc_bldg_flooded_max,
    fill = F,
    color = "red",
    weight = 0.5
  ) %>% 
  addLegend(
    pal = flood_pal,
    values = values(flood_max),
    title = "Flood depth, cm"
  )

```

```{r}
fc_bldg_exposure <- NULL

# for(slr in c("000","025","050")){
#   
#   for(rp in c("001","020","100")){
#     
#     print(paste0("SLR",slr,"_RP",rp))
#     
#     flood <- raster( paste0("/Users/alessandrokerr/Documents/agkerr99.github.io/218y/Assignment 4/flood/SLR",slr,"_RP",rp,"_fc_flood.tif"))
#     
#     flood_extent <- 
#       (flood > -Inf) %>% 
#       st_as_stars() %>% 
#       st_as_sf(merge = T) %>% 
#       st_set_crs(26910) %>% 
#       st_make_valid() %>% 
#       st_transform(4326)
#     
#     fc_bldg_flooded <-
#       fc_bldg_flooded_max[flood_extent,] %>% 
#       st_transform(26910)
#   
#     if(nrow(fc_bldg_flooded) == 0) {break}
#       
#     flood_crop <-
#       crop(flood, fc_bldg_flooded)
#     
#     flood_crop[is.na(flood_crop)] <- 0
#     
#     temp <-
#       raster::extract(
#         flood_crop,
#         fc_bldg_flooded,
#         fun = mean
#       ) %>% 
#       as.data.frame() %>% 
#       rename(avg_depth = V1) %>% 
#       cbind(
#         fc_bldg_flooded %>% 
#           st_drop_geometry() %>% 
#           dplyr::select(osm_id)
#       ) %>% 
#       mutate(
#         SLR = slr,
#         RP = rp
#       )
#     
#     fc_bldg_exposure <- 
#       fc_bldg_exposure %>% 
#       rbind(temp)
#     
#   }
# }
# saveRDS(fc_bldg_exposure,"fc_bldg_exposure.rds")


```

```{r}
sm_cbg <- block_groups("CA","San Mateo", cb = F, progress_bar = F, year = 2019)

fc_cbg<-sm_cbg %>%
  st_centroid() %>%
  .[zip,] %>%
  st_set_geometry(NULL)%>%
  left_join(
    sm_cbg %>% select(GEOID)
    ) %>%
  st_as_sf()
  
```

```{r}
fc_block_veh_per_bldg <-
  fc_bldg %>% 
  filter(type == "residential" | type == "house" | is.na(type)) %>% # any non-residential buildings?
  select(osm_id) %>% # unique ID for each building
  st_centroid() %>% 
  st_join(zip_blocks %>% select(GEOID20)) %>% # block shapes
  st_join(sm_cbg %>% select(cbg = GEOID)) %>% # cbg shapes
  st_drop_geometry() %>% 
  group_by(cbg, GEOID20) %>% # "assuming population is distributed evenly across buildings in a block"
  summarize(bldg_count = n()) %>% # how to get counts?
  left_join(fc_pop_2020, by = c("GEOID20" = "block")) %>%
  rename("pop" = "estimate") %>% # census dataset
  left_join(fc_vehicles, by = c("cbg")) %>%
  rename("vehicle" = "estimate") %>% # census dataset
  filter(!is.na(pop)) %>%
  filter(!is.na(vehicle)) %>%
  filter(pop != 0) %>%
  group_by(cbg) %>% # "and vehicles are distributed evenly across population"
  mutate(
    veh_per_person = vehicle/sum(pop),
    ppl_per_bldg = pop/bldg_count,
    veh_per_bldg = veh_per_person*ppl_per_bldg # fractional result ok
  )


```
# PART 4: Vulnerability Data

```{r}
vulnerability <- data.frame(
  depth_above_ground = c(0.5,1:10),
  perc_damage = c(
    0.076,
    0.28,
    0.462,
    0.622,
    0.76,
    0.876,
    0.97,
    1,
    1,
    1,
    1
  )
) %>%
  mutate(
    depth_cm = depth_above_ground*30.48
  )

library(ggplot2)

ggplot(vulnerability, aes(depth_above_ground, perc_damage)) +
  geom_point(shape = 16, size = 5) +
  theme_minimal() + 
  geom_line()
  
```
** ASSUMPTION: SEDANS
```{r}
# fc_bldg_exposure <-
#   readRDS("fc_bldg_exposure.rds")
# 
# fc_veh_perc_damage <-
#   approx(
#     x = vulnerability$depth_cm,
#     y = vulnerability$perc_damage,
#     xout = fc_bldg_exposure$avg_depth
#   ) %>%
#   .[2] %>%
#   as.data.frame() %>%
#   rename(perc_damage = y) %>%
#   cbind(fc_bldg_exposure) %>%
#   mutate(
#     perc_damage = ifelse(
#       is.na(perc_damage),
#       0,
#       perc_damage
#     ))
# 
# saveRDS(fc_veh_perc_damage,"fc_veh_perc_damage.rds")

fc_veh_perc_damage <- readRDS("fc_veh_perc_damage.rds")
```

```{r}
fc_veh_perc_damage_plot <- 
  expand.grid(
    osm_id = unique(fc_veh_perc_damage$osm_id),
    SLR = unique(fc_veh_perc_damage$SLR),
    RP = unique(fc_veh_perc_damage$RP)
  ) %>% 
  left_join(fc_veh_perc_damage) %>% 
  mutate(
    perc_damage = ifelse(
      is.na(perc_damage),
      0,
      perc_damage
    ))
```


```{r}
library(plotly)

fc_plot <- 
  plot_ly() %>% 
  add_trace(
    data = 
      fc_veh_perc_damage_plot %>% 
        filter(RP == "100") %>% 
        mutate(SLR = SLR %>% as.numeric()),
    x = ~avg_depth,
    y = ~perc_damage,
    frame = ~SLR,
    type = 'scatter',
    mode = 'markers',
    marker = list(
      color = 'rgba(17, 157, 255, 0.01)',
      size = 15
    ),
    showlegend = F
  ) %>% 
  add_trace(
    data = vulnerability,
    x = ~depth_cm,
    y = ~perc_damage,
    type = 'scatter',
    mode = 'markers',
    marker = list(
      color = 'rgb(0,0,0)'
    ),
    showlegend = F
  ) %>% 
  layout(
    xaxis = list(
      title = "Average Flood Depth",
      zeroline = FALSE
    ),
    yaxis = list(
      title = "Percent Damage"
    ),
    title = "Foster City building damage during<br>100-year storm, by base sea level rise"
  ) %>% 
  config(displayModeBar = F)

fc_plot
```


# PART 5: RISK ESTIMATION

```{r}
#detach("package:raster")

fc_veh_damage <-
  fc_veh_perc_damage %>%  
  left_join(fc_bldg %>% select(osm_id)) %>%
  st_as_sf() %>%
  st_centroid() %>%
  st_join(fc_block_veh_per_bldg %>% st_as_sf() %>% select(veh_per_bldg)) %>%
  #filter(veh_per_bldg < 4) %>%
  mutate(
    damage = perc_damage * veh_per_bldg * 27000 * (1-0.505)
  ) %>% 
  select(osm_id, SLR, RP, damage) %>%
  filter(!is.na(damage))
```
Kelley Blue Book value = 27,000 (average used car price)
Assume: None are immune to flooding
% of people who move with advanced warning: 6 hours or less advanced warning = 50.5% https://planning.erdc.dren.mil/toolbox/library/EGMs/egm09-04.pdf
Assume: if vehicles per building >4 they're likely not single family residential homes?

```{r}
fc_veh_aal_by_slr <-
  fc_veh_damage %>% 
  st_drop_geometry() %>%
  pivot_wider(
    names_from = RP,
    values_from = damage
  ) %>% 
  replace(is.na(.), 0) %>% 
  mutate(
    damage = 
      0.95*(`001`+`020`)/2 + 
      0.04*(`020`+`100`)/2 + 
      0.01*(`100`)
  ) %>% 
  select(osm_id, SLR, damage)
```

```{r}
rcp45 <- read_csv("https://raw.githubusercontent.com/stanfordfuturebay/stanfordfuturebay.github.io/master/advanced/rcp45_sanfrancisco.csv")

rcp45
```

```{r}
fc_veh_aal_by_year <- 
  fc_veh_aal_by_slr %>% 
  left_join(
    rcp45 %>% 
      mutate(
        SLR = str_pad(SLR, 3 , "left", "0")
      ) %>% 
      select(
        SLR,
        `2020`,
        `2030`,
        `2040`,
        `2050`
      )
  ) %>% 
  pivot_longer(
    `2020`:`2050`,
    names_to = "year",
    values_to = "occurrence"
  ) %>% 
  pivot_longer(
    c(damage,occurrence),
    names_to = "key",
    values_to = "value"
  ) %>% 
  pivot_wider(
    names_from = c("key","SLR"),
    values_from = value
  ) %>% 
  replace(is.na(.), 0) %>% 
  mutate(
    damage = 
      occurrence_050 * (damage_050)
  ) %>% 
  mutate(
    `Calendar Year` = as.numeric(year)
  ) %>%
  select(osm_id, `Calendar Year`, damage) %>%
  left_join(emfac, by = "Calendar Year") %>%
  mutate(
    Increase =
  ifelse(is.na(Increase), 0, Increase)) %>%
  mutate(
    projected_damage = damage*(1+Increase/100)) %>%
  select(osm_id, `Calendar Year`, projected_damage) %>%
  rename("year" = "Calendar Year") %>%
  rename("damage" = "projected_damage")
```
```{r}
fc_veh_aal_by_year_map <-
  fc_veh_aal_by_year %>% 
  pivot_wider(
    names_from = year,
    values_from = damage
  ) %>% 
  mutate(
    change = `2050`-`2020`
  ) %>% 
  left_join(
    fc_bldg_flooded_max %>%
      select(osm_id)
  ) %>% 
  st_as_sf() %>% 
  st_transform(4326)
```

```{r}
aal_pal <- colorNumeric(
  palette = "Reds",
  domain = c(0,fc_veh_aal_by_year_map$`2050`)
)

fc_veh_aal_by_year_map %>% 
  leaflet() %>% 
  addMapboxTiles(
    style_id = "light-v9",
    username = "mapbox"
  ) %>% 
  addPolygons(
    fillColor = ~aal_pal(`2020`),
    color = "gray",
    fillOpacity = 1,
    opacity = 1,
    weight = 0.25,
    highlightOptions = highlightOptions(
      color = "white",
      weight = 2
    ),
    label = ~paste0("$",prettyNum(signif(`2020`,2),",")," average annualized loss in 2020"),
    group = "2020"
  ) %>% 
  addPolygons(
    fillColor = ~aal_pal(`2050`),
    color = "gray",
    fillOpacity = 1,
    opacity = 1,
    weight = 0.25,
    highlightOptions = highlightOptions(
      color = "white",
      weight = 2
    ),
    label = ~paste0("$",prettyNum(signif(`2050`,2),",")," average annualized loss in 2050"),
    group = "2050"
  ) %>% 
  addPolygons(
    fillColor = ~aal_pal(change),
    color = "gray",
    fillOpacity = 1,
    opacity = 1,
    weight = 0.25,
    highlightOptions = highlightOptions(
      color = "white",
      weight = 2
    ),
    label = ~paste0("$",prettyNum(signif(change,2),",")," change in average annualized loss from 2020 to 2050"),
    group = "Change"
  ) %>% 
  addLegend(
    pal = aal_pal,
    values = ~`2050`,
    title = "AAL"
  ) %>% 
  addLayersControl(
    baseGroups = c("2020","2050","Change"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  showGroup("2050")
```
Make comment about remaining apartments etc. Dense population in apartment blocks skews the vehicle counts per building in the block. 

```{r}
fc_boundary <- 
  places("CA", cb = T, progress_bar = F) %>% 
  filter(NAME == "Foster City")

fc_bgs <- 
  block_groups("CA", "San Mateo", progress_bar = F) %>% 
  select(GEOID) %>% 
  .[fc_boundary, ]
```

```{r}
fc_bg_aal <-
  fc_veh_aal_by_year %>% 
  pivot_wider(
    names_from = year,
    values_from = damage
  ) %>% 
  mutate(
    aal = (`2020`*5 + `2030`*10 + `2040`*10 + `2050`*5)/30
  ) %>% 
  left_join(
    fc_bldg_flooded_max %>%
      select(osm_id) %>% 
      st_centroid()
  ) %>% 
  st_as_sf() %>% 
  st_transform(4269) %>% 
  st_join(fc_bgs) %>% 
  st_set_geometry(NULL) %>% 
  group_by(GEOID) %>% 
  summarize(
    aal = sum(aal),
    count = n()
  ) %>% 
  left_join(fc_bgs) %>% 
  st_as_sf()
```
```{r}
aal_pal <- colorNumeric(
  palette = "Reds",
  domain = fc_bg_aal$aal
)

fc_bg_aal %>% 
  leaflet() %>% 
  addMapboxTiles(
    style_id = "light-v9",
    username = "mapbox"
  ) %>% 
  addPolygons(
    fillColor = ~aal_pal(aal),
    color = "gray",
    fillOpacity = 0.5,
    opacity = 1,
    weight = 0.25,
    highlightOptions = highlightOptions(
      color = "white",
      weight = 2
    ),
    label = ~paste0("$",prettyNum(signif(aal,2),",")," average annualized loss across ", count, " buildings, 2020-2050")
  ) %>% 
  addLegend(
    pal = aal_pal,
    values = ~aal,
    title = "AAL, 2020-2050"
  )
```

```{r}
fc_flood_risk <- fc_no_one_vehicle %>%
  ungroup() %>%
  mutate(sum_one_no_veh = sum(estimate)) %>%
  mutate(perc_risk = sum_one_no_veh/nrow(fc_bldg_flooded))

paste0("Number of Households at Flood Risk in Foster City with One or No Vehicles: ",round(fc_flood_risk$perc_risk[1]*100,0), " %")
```
If you have only one or no vehicles, you lose access to daily activities (work, etc). Odds are lower that you can remove your vehicle in time if you only have one vehicle. For households with no vehicles, you are stuck there duirng the flood. Public transit would likely not be running and you would be somewhat trapped at your houes. 


