---
title: "Assignment 6 - Resumbmission"
author: "Alessandro Kerr"
date: " Nov 29, 2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F, message = F)

library(tidyverse)
library(plotly)
library(sf)
library(tigris)
library(leaflet)
library(censusapi)
library(lubridate)
library(dplyr)
library(mapview)
```

#TASK 1

Firstly, the PUMA's data will be filtered down to only San Francisco. There are three strange geographic anamolies that should be noted, but should not affect the analysis. Two instances in group 6, and one instance in group 2 appear to be small geographic areas that do not correlate with the rest of the group. These are due to islands that are technically a part of San Francisco. See the plotted PUMAs below.

```{r}
Sys.setenv(CENSUS_KEY="0f0e2bffc9d96d5405bcb932e25318a4edb32602")

```

```{r}
ca_pumas <- pumas("CA", cb = T, progress_bar = F)

sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()
```

```{r}
mapview(sf_pumas)
```

#TASK 2

Now, the PUMs data for the PUMAs was retrieved. 

```{r}
pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "NP",
    "HHL",
    "HINCP",
    "TEN",
    "MV",
    "YBL",
    "BLD",
    "AGEP"
  )
)   
```

```{r}
sf_pums <- bay_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% sf_pumas$PUMACE10)
```

#PART 3

From here, structures were filtered for those built before 1960. This is a simplification of the Lead Risk Chapter of the CES report.

```{r}
sf_house_1960 <- sf_pums %>%
  filter(YBL == c(1,2,3))
```

#PART 4

Now, the leadrisk variable was created based on household income being less than $90k, and at least one child being less than 6 years old.

```{r}
sf_pums_clean <- sf_pums %>%
  mutate(
    YBL = as.numeric(YBL),
    AGEP = as.numeric(AGEP),
    as.numeric(HINCP)
  ) %>%
  filter(YBL %in% c(1:3)) %>%
  arrange(AGEP) %>%
  group_by(SERIALNO) %>%
  summarize_all(first) %>%
  mutate(leadrisk = ifelse(
    as.numeric(HINCP) < 90000 & as.numeric(AGEP) < 6,
    1,
    0
  ))
```

#PART 5

Next, a logit model was created for leadrisk based on the BLD, TEN, MV, and PUMA variables. The summary of the model can be seen below. 

```{r}
sf_pums_factors <- sf_pums_clean %>%
  mutate(
    pumas = public_use_microdata_area %>%
      factor(
        levels = sf_pums_clean$PUMA %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ),
    units = BLD %>%
      factor(
        levels = sf_pums_clean$BLD %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ),
    tenure = TEN %>%
      factor(
        levels = sf_pums_clean$TEN %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ),
    move = MV %>%
      factor(
        levels = sf_pums_clean$MV %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )
  )
```

```{r}

model <- glm(
  leadrisk ~ units + pumas + tenure + move,
  family = quasibinomial(),
  data = sf_pums_factors
)


summary(model)
```


Additionally, a predicted liklihood score can be seen for a random row in the dataset.


```{r}
random <- sample_n(sf_pums_factors,1)
random
predict(model, data.frame(random), type = "response")
```

#PART 6

```{r}
full_predict <- predict(model, sf_pums_factors, type = "response" )

sf_pums_predict <- sf_pums_factors %>%
  cbind(full_predict)

```

```{r}
predicted1 <-
  sf_pums_predict %>% 
  mutate(
    leadrisk = ifelse(
      leadrisk == 1, 
      "Yes (Leadrisk)", 
      "No (Leadrisk)")) %>%
  mutate(
    pred_yes = ifelse(
        full_predict >0.1, WGTP, 0)) %>%
  mutate(
      pred_no = ifelse(
        full_predict < 0.1, WGTP, 0)) %>%
  group_by(leadrisk) %>%
  summarize(pred_yes = sum(as.numeric(pred_yes)), pred_no = sum(as.numeric(pred_no)))

predicted1

```

The following can be stated about the summary of the model. There were 421 cases of leadrisk that the model accurately predicted. There were 221620 cases where the model correctly predicted there is no lead risk. Conversely, there were 2777 cases where households did not have a lead risk, but the model predicted they did. Lastly, there were 3988 cases where the model predicted no lead risk, but they actually did have one. 

In simple heuristics, the households that typically exceeded the threshold were single family houses,  mostly rented or occupied without rent, and had been moved into 2-10 years ago. 

Based on their threshold score of 10%, they would mail out 3198 packages. Based on the model, it is estimated that 421/3198, or 13% of postcards would actually go to low-income households with children. In the raw data, there were 4409 households with a leadrisk. Therefore, 421/4409, or 9.5% of all at-risk children are likely to be reached. 

#PART 7

```{r}
temp <- tempfile()
download.file("https://www2.census.gov/programs-surveys/acs/data/pums/2019/1-Year/csv_hca.zip",destfile = temp)
pums_hca_2019_1yr <- read_csv(unzip(temp,"psam_h06.csv"))
unlink(temp)

```

```{r}
pums_2019_1yr_wts <- pums_hca_2019_1yr %>%
  filter(PUMA %in% sf_pumas$PUMACE10) %>%
  mutate(YBL = as.numeric(YBL)) %>%
  filter(YBL %in% c(1:3))

sf_pums_wts <- sf_pums_predict %>%
  left_join(pums_2019_1yr_wts,
            by = "SERIALNO") %>%
  select(starts_with("WGTP"), -WGTP.x, -WGTP.y)


```

```{r}
library(survey)

logit_survey_design <- svrepdesign(
  data = sf_pums_predict,
  type = "ACS",
  repweights = sf_pums_wts,
  weights = ~as.numeric(WGTP)
)
```

```{r}
logit_survey_model <- svyglm(
  formula = leadrisk ~ units + pumas + tenure + move,
  family = quasibinomial(),
  design = logit_survey_design
)

summary(logit_survey_model)
```

```{r}
#exp(coef(logit_survey_model))
```

```{r}
predict(logit_survey_model, data.frame(random), type = "response")
```

The weighted model predicted a similar value to that of the linear model. 

```{r}

logit_predict <- predict(logit_survey_model, sf_pums_factors, type = "response")

sf_pums_logit <- sf_pums_factors %>%
  cbind(logit_predict)

```

```{r}
predicted2 <-
  sf_pums_logit %>% 
  mutate(
    leadrisk = ifelse(
      leadrisk == 1, 
      "Yes (Leadrisk)", 
      "No (Leadrisk)")) %>%
  mutate(
    pred_yes = ifelse(
        response >0.1, WGTP, 0)) %>%
  mutate(
      pred_no = ifelse(
        response < 0.1, WGTP, 0)) %>%
  group_by(leadrisk) %>%
  summarize(pred_yes = sum(as.numeric(pred_yes)), pred_no = sum(as.numeric(pred_no)))

predicted2
```

There are some key differences between the previous model and the weighted model. Now, there were 1835 cases of leadrisk that the model accurately predicted. There were 214,964 cases where the model correctly predicted there is no lead risk. Conversely, there were 9433 cases where households did not have a lead risk, but the model predicted they did. Lastly, there were 2574 cases where the model predicted no lead risk, but they actually did have one. 

Based on their threshold score of 10%, they would mail out 11,268 packages. Based on the model, it is estimated that 1835/11268, or 16.2% of postcards would actually go to low-income households with children. This is a significant increase over the previous unweighted scenario. 